Finally Tagless, Partially Evaluated

### 2008-10-23 tagless/tagless-final.tex
Refer to Mogensen-SelfApplicable, which uses the same fold encoding as we do, but untyped

We emphasize ``typed'' and ``fold'' in the previous sentence.
Our encoding of the recursive type of terms \citep{bohm-automatic}
makes folds over terms much easier to write than functions over
terms that are not primitive recursive (or, compositional).
In contrast, \citearound{'s encoding of the sum type of
terms}\citet{JFP-Mogensen} does not privilege folds.
In exchange, we statically express object types in the metalanguage
and prevent both kinds of run-time errors in
\S\ref{tagproblem}, due to evaluating ill-typed or open terms.
Because the new interpreter
uses no universal type or pattern matching, it never gives a
run-time error, and is in fact total.  Because this safety is obvious
not just to us but also to the metalanguage implementation, we avoid
the serious performance penalty \citep{WalidICFP02}
that arises from error checking at run time.

@@

Our partial evaluator owes much to
\citet{Thiemann-combinators} and \citet{sumii-hybrid}, who deforested
the object term representation and expressed a partial evaluator as a collection
of term combinators in a typed metalanguage.  Like us, \citeauthor{sumii-hybrid}
follow \citet{Sperber-SelfApplicable} and \citet{asai-binding-time} in building
static and dynamic terms in tandem, to combine offline and online partial
evaluation.
\citearound{'s earlier self-reducer for the untyped
$\lambda$-calculus}\Citet{JFP-Mogensen} also
builds static and dynamic terms in tandem.  However, it
builds a static term for every object term, even a bound variable,
so it moves some work from |app| to~|pdyn| (in terms of Figure~\ref{fig:pe})
and remains untyped.  In contrast, we follow Sperber, Asai,
and \citeauthor{sumii-hybrid} in leaving the static term
optional, so as to perform lifting without
juggling explicit type indices in the encoding of an object term.

>>>>

We emphasize ``typed'' and ``fold'' in the previous sentence.
We use a typed version of \citets{Mogensen-SelfApplicable}
encoding of the recursive type of terms \citep{bohm-automatic}, which
makes folds over terms much easier to write than functions over
terms that are not primitive recursive (or, compositional).
In contrast, \citearound{'s earlier encoding of the sum type of
terms}\citet{JFP-Mogensen} does not privilege folds.
In exchange, we statically express object types in the metalanguage
and prevent both kinds of run-time errors in
\S\ref{tagproblem}, due to evaluating ill-typed or open terms.
Because the new interpreter
uses no universal type or pattern matching, it never gives a
run-time error, and is in fact total.  Because this safety is obvious
not just to us but also to the metalanguage implementation, we avoid
the serious performance penalty \citep{WalidICFP02}
that arises from error checking at run time.

@@

Our partial evaluator owes much to
\citet{Thiemann-combinators} and \citet{sumii-hybrid}, who deforested
the object term representation and expressed a partial evaluator as a collection
of term combinators in a typed metalanguage.  Like us, \citeauthor{sumii-hybrid}
follow \citet{Sperber-SelfApplicable} and \citet{asai-binding-time} in building
static and dynamic terms in tandem, to combine offline and online partial
evaluation.
\citearound{'s earlier self-reducers for the untyped
$\lambda$-calculus}\Citet{JFP-Mogensen,Mogensen-SelfApplicable} also
build static and dynamic terms in tandem.  However, they
build a static term for every object term, even a bound variable,
so they move some work from |app| to~|pdyn| (in terms of Figure~\ref{fig:pe})
and remain untyped.  In contrast, we follow Sperber, Asai,
and \citeauthor{sumii-hybrid} in leaving the static term
optional, so as to perform lifting without
juggling explicit type indices in the encoding of an object term.
